\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2021-07-11 manuscript.Rnw
Rename Section 4 to 'Simulation of model outcomes'

The remainder of this paper is organized as follows. Section~\ref{sec:model-taxonomy} provides a mathematical description of the economic models supported by \pkg{hesim}. An overview of the coding framework is provided in Section~\ref{sec:framework}. An illustrative example using a three state model of disease progression in oncology is provided in Section~\ref{sec:illustrative-example}. Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient-level and aggregate-level data. The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples based on the output from the prior section. Section~\ref{sec:discussion} discusses additional features not covered in this paper, makes comparisons to other software, and explores possible extensions. Finally, Section~\ref{sec:conclusion} concludes.

@@

where $\Delta t_c = t_{c} - t_{c-1}$,  $t_c^{*}\in[t_{c-1},t_{c}]$, and $f(t)= z_h(t) e^{-rt} p_h(t)$. Three methods can be used to estimate $f(t_c^{*})$. First, a left Riemann sum uses values at the start of each time interval, $f(t_c^{*}) = f(t_{c-1})$. Second, a right Riemann sum uses values at the end of each time interval, $f(t_c^{*}) = f(t_{c})$. Finally, the trapezoid rule averages values at the start and end of each interval, $f(t_c^{*}) = \frac{1}{2}\Delta t_c[f(t_{c-1}) + f(t_{c})]$.

@@

\begin{table} [h]
\footnotesize
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
\hline
Economic model & Disease progression  & QALYs & Costs \\
\hline
\code{CohortDtstm} & \code{sim_stateprobs()} & \code{sim_qalys()} & \code{sim_costs()}\\
\code{IndivCtstm} & \code{sim_disease()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
\code{Psm} & \code{sim_survival()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
\hline
\end{tabular*}
\caption{Methods for simulating outcomes from economic models.}\label{tbl:simulate-outcomes}
\end{table}

%% -- Illustrative example -----------------------------------------------------
\section{Illustrative example} \label{sec:illustrative-example}
We consider a three-state model commonly used in oncology. As shown in Figure~\ref{fig:three-state-model}, the three health states are stable disease (i.e., not progressed), progression, and death. Patients can transition to a more severe health state (stable $\rightarrow$ progression, stable $\rightarrow$ death, and progression $\rightarrow$ death) but cannot recover to a less severe health state. We assume that patients remain on first line (1L) treatment until progression, at which time they switch to a second line (2L) treatment.

@@

\subsection{Application}
Uncertainty over $\theta$ is captured using the PSA. The simulated distribution of QALYs and costs for each treatment strategy can be produced with the \code{$summarize()} method as was demonstrated in Section~\ref{sec:illustrative-example}. We will use the output from the iCTSTM in this example.

@@

%% -- Discussion ---------------------------------------------------------------
\section{Discussion} \label{sec:discussion}
This paper formally describes the economic models and cost-effectiveness framework used by the \pkg{hesim} package. An illustrative example is provided for a three state model commonly used to evaluate treatments in oncology. However, since \pkg{hesim} aims to be flexible, the example only covers a subset of the relevant modeling approaches and functionality. Additional examples as well as documentation for all functions and classes are provided on the package website (\url{https://hesim-dev.github.io/hesim/}).

>>>>

The remainder of this paper is organized as follows. Section~\ref{sec:model-taxonomy} provides a mathematical description of the economic models supported by \pkg{hesim}. An overview of the coding framework is provided in Section~\ref{sec:framework}. Section~\ref{sec:model-simulation} uses a three state model of disease progression in oncology to show how \pkg{hesim} can be used to simulate model outcomes. Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient-level and aggregate-level data. The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples based on the output from the prior section. Section~\ref{sec:discussion} discusses additional features not covered in this paper, makes comparisons to other software, and explores possible extensions. Finally, Section~\ref{sec:conclusion} concludes.

@@

where $\Delta t_c = t_{c} - t_{c-1}$,  $t_c^{*}\in[t_{c-1},t_{c}]$, and $f(t)= z_h(t) e^{-rt} p_h(t)$. Three methods can be used to estimate $f(t_c^{*})$. First, a left Riemann sum uses values at the start of each time interval, $f(t_c^{*}) = f(t_{c-1})$. Second, a right Riemann sum uses values at the end of each time interval, $f(t_c^{*}) = f(t_{c})$. Finally, the trapezoid rule averages values at the start and end of each interval, $f(t_c^{*}) = \frac{1}{2}[f(t_{c-1}) + f(t_{c})]$.

@@

\begin{table} [h]
\caption{Methods for simulating outcomes from economic models.}\label{tbl:simulate-outcomes}
\footnotesize
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
\hline
Economic model & Disease progression  & QALYs & Costs \\
\hline
\code{CohortDtstm} & \code{sim_stateprobs()} & \code{sim_qalys()} & \code{sim_costs()}\\
\code{IndivCtstm} & \code{sim_disease()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
\code{Psm} & \code{sim_survival()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
\hline
\end{tabular*}
\end{table}

%% -- Illustrative example -----------------------------------------------------
\section{Simulation of model outcomes} \label{sec:model-simulation}
To demonstrate how \pkg{hesim} can be used to simulate outcomes from an economic model, we consider a three state model commonly used in oncology. As shown in Figure~\ref{fig:three-state-model}, the three health states are stable disease (i.e., not progressed), progression, and death. Patients can transition to a more severe health state (stable $\rightarrow$ progression, stable $\rightarrow$ death, and progression $\rightarrow$ death) but cannot recover to a less severe health state. We assume that patients remain on first line (1L) treatment until progression, at which time they switch to a second line (2L) treatment.

@@

\subsection{Application}
Uncertainty over $\theta$ is captured using the PSA. The simulated distribution of QALYs and costs for each treatment strategy can be produced with the \code{$summarize()} method as was demonstrated in Section~\ref{sec:model-simulation}. We will use the output from the iCTSTM in this example.

@@

%% -- Discussion ---------------------------------------------------------------
\section{Discussion} \label{sec:discussion}
This paper formally describes the economic models and cost-effectiveness framework used by the \pkg{hesim} package. An illustrative example was provided for a three state model commonly used to evaluate treatments in oncology. However, since \pkg{hesim} aims to be flexible, the example only covers a subset of the relevant modeling approaches and functionality. Additional examples as well as documentation for all functions and classes are provided on the package website (\url{https://hesim-dev.github.io/hesim/}).
